Mixture-of-tastes Models for Representing Users with Diverse Interests

### 2018-01-29 paper/paper.tex

\citet{athiwaratkun2017multimodal} extend this idea by representing words a mixtures of Gaussians. This allows them to capture polysemy by modelling each word with several distinct, small-variance Gaussian distributions (rather than artificially inflating the variance of a single distribution). This allows much clearer separation of distinct meanings. This work is closest to the approach presented here.

@@

@@

The second factorization model we test is the Projection-Mixture Factorization (\textbf{PM-F}) model. It extends the basic matrix factorization model by using linear projection layers in ocnjunction with latent vectors to model multiple tastes: instead of embedding $m$ taste vectors and $m$ mixture vectors for each user directly, we use $2m$ $k \times k$ projection matrices to project a single $k$-dimensional user item vector into $m$ taste and $m$ mixture vectors. These matrices are shared between all users, making the model substantially more parsimonious.

>>>>

\citet{athiwaratkun2017multimodal} extend this idea by representing words as mixtures of Gaussians. This allows them to capture polysemy by modelling each word with several distinct, small-variance Gaussian distributions (rather than artificially inflating the variance of a single distribution). This allows much clearer separation of distinct meanings.

@@

The approach closest to the one presented here, but focusing solely on the sequence-based setting and using pre-trained item embeddings, is presented in \citet{wang2017attention}, and was published contemporaneously with the writing of this paper. The authors use recurrent neural networks with attention to model user sequences, and either feedforward or recurrent decoders to generate taste mixture components. Their experiments confirm that the addition of taste mixtures improves recommendation quality. We suggest that the two papers should be seen as mutually reinforcing and complementary, and that a combination of both approaches (recurrent decoders and end-to-end embedding training) would be beneficial.

@@

The second factorization model we test is the Projection-Mixture Factorization (\textbf{PM-F}) model. It extends the basic matrix factorization model by using linear projection layers in conjunction with latent vectors to model multiple tastes: instead of embedding $m$ taste vectors and $m$ mixture vectors for each user directly, we use $2m$ $k \times k$ projection matrices to project a single $k$-dimensional user item vector into $m$ taste and $m$ mixture vectors. These matrices are shared between all users, making the model substantially more parsimonious.
